---
title: "Case Study 2"
author: "Panayiota Harken Alexiou"
date: "2023-01-02"
mainfont: Times New Roman
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Calculations ##

### Total revenue per year ###

```{r, warning=FALSE,message=FALSE}
data=read.csv("casestudy.csv")
dim(data)

```
### Total revenue per year ###
```{r, warning=FALSE,message=FALSE}
library(tidyverse)
library(ggthemes)

year <- data %>%
  select(year) %>%
  unique()
result <- tibble(year)
result$year <- as.integer(result$year)

total_revenue <- data %>%
  group_by(year) %>%
  mutate(total_revenue = sum(net_revenue)) %>%
  select(total_revenue, year) %>%
  unique()


result=result %>% left_join(total_revenue, by="year")
result


```
### New Customer Revenue ###

```{r, message=FALSE,warning=FALSE}

data <- data %>%
  group_by(customer_email) %>%
  mutate(new_customer = case_when(year == min(year) ~ TRUE, year > min(year) ~ FALSE))

new_customer_revenue <- data %>%
  filter(new_customer == TRUE) %>%
  group_by(year) %>%
  mutate(new_customer_revenue = sum(net_revenue)) %>%
  select(new_customer_revenue, year) %>%
  unique()

result <- result %>%
  left_join(new_customer_revenue, by = "year")

result
```
### Existing Customer Growth ###


```{r, message=FALSE,warning=FALSE}

result <- result %>%
  arrange(year) %>%
  mutate(existing_customer_growth = (total_revenue - new_customer_revenue) - (lag(total_revenue) - lag(new_customer_revenue)))

result

```
### Revenue lost from attrition ###


```{r, message=FALSE,warning=FALSE}
result <- result %>%
  arrange(year) %>%
  mutate(revenue_lost_attrition = lag(total_revenue) - (total_revenue - new_customer_revenue))

result
```
### Existing customer output current year###


```{r, message=FALSE,warning=FALSE}

result <- result %>%
  mutate(existing_customer_revenue = total_revenue - new_customer_revenue)

result
```

### Existing customer output prior year###


```{r, message=FALSE,warning=FALSE}

result <- result %>%
 arrange(year) %>%
  mutate(existing_customer_revenue_prior = lag(total_revenue) - lag(new_customer_revenue))

result
```

### Total Customers Current and Previous year ###


```{r, message=FALSE,warning=FALSE}
total_customers <- data%>%
  group_by(year) %>%
  mutate(total_customers = n()) %>%
  select(total_customers, year) %>%
  unique() %>%
  ungroup() %>%
  arrange(year) %>%
  mutate(total_customers_prior = lag(total_customers))

result <-result %>%
  left_join(total_customers, by = "year")

result

```

### New and Lost Customers ###


```{r, message=FALSE,warning=FALSE}

new_customers <- data %>%
  filter(new_customer == TRUE) %>%
  group_by(year) %>%
  mutate(new_customers = n()) %>%
  select(new_customers, year) %>%
  unique()

result <- result %>%
  left_join(new_customers, by = "year")

result <- result %>%
  arrange(year) %>%
  mutate(lost_customers = lag(total_customers) - (total_customers - new_customers))

result 
```

## Visualization ##


```{r, message=FALSE,warning=FALSE}

ggplot(result, aes(x = factor(year), y = total_revenue)) + geom_bar(stat = "identity")

```
<br>
It is obvious that on 2017 the total revenue is higher than the other years. The plot of new customers shows that new customers are more on 2017. 

```{r, message=FALSE,warning=FALSE}

ggplot(result, aes(x = factor(year), y = new_customers)) + geom_bar(stat = "identity")

```
<br>
From the following plot it is obvious that the revenue is higher from the existing customers and not from the new ones on average. 

```{r, message=FALSE,warning=FALSE}
result %>%
  ggplot() +
  geom_line(aes(x = year, y = new_customer_revenue/new_customers, color = "New")) +
  geom_line(aes(x = year, y = (total_revenue - new_customer_revenue)/ (total_customers - new_customers),  color = "Existing")) +
  labs(x = "Year", y = "Average Revenue per Customer", title = "Average Revenue for New and Existing Customers") +
  scale_color_manual("", breaks = c("Existing", "New"), values = c("Existing" = "magenta", "New" = "yellow2"))


```
